% !TeX spellcheck = en_US
\section{Diffusion Map\label{Sec:DR:DM}}
A diffusion map was developed by Coifman and Lafon in 2006\cite{CoifmanACHA2006}, which embeds data in a lower-dimensional space, such that the diffusion distance in the original space between points is well approximated by the Euclidean distance in the reduced-dimensional space.

Given a set of $N$ data points $\mathbf{x}=\{x_1, x_2,\dots,x_N\}$, the connectivity between data points $x_i$ and $x_j$ is defined by the transition probability between these two points $p(x_i,x_j)$, which is measured by their distance
\begin{equation}
	p(x_i,x_j)\propto \begin{cases}
		k(x_i,x_j), & \text{if } k(x_i,x_j) >\epsilon\\
		0, & \text{otherwise}
	\end{cases},
\end{equation}
with $0<\epsilon \ll 1$. The kernel $k(x,y)$ defines a local measure of similarity within a certain neighborhood, and Gaussian kernel
\begin{equation}
	k(x,y)=\exp{\left(-\frac{|x-y|^2}{\alpha}\right)}
\end{equation}
is frequently used, where $\alpha$ tunes the size of the neighborhood. Transition probability matrix $\mathbf{P}$ must be row-normalized, which leads to
\begin{equation}
	p(x_i,x_j)= \begin{cases}
		k(x_i,x_j)/\sum\limits_{x_j\in S_i}k(x_i,x_j), & \text{if } k(x_i,x_j) >\epsilon\\
		0, & \text{otherwise}
	\end{cases}.
\end{equation}
$S_i$ is the neighborhood of $x_i$.
 
Suppose there are three data points $\{x_1, x_2, x_3\}$ and the single-step transition probability matrix is
\begin{equation}
  P=\begin{bmatrix}
	p_{11} & p_{12} & 0\\
	p_{21} & p_{22} & p_{23} \\
	0      & p_{32} & p_{33}
\end{bmatrix},
\end{equation}
where we have assumed that the transition probability from $x_1$ to $x_3$ is zero and vice versa. It can be easily found that
\begin{equation}
  P^2=
  \begin{bmatrix}
  	p_{11}p_{11}+p_{12}p_{21} & p_{11}p_{12}+p_{12}p_{22} & p_{12}p_{23}\\
  	p_{21}p_{11}+p_{22}p_{21} & p_{21}p_{12}+p_{22}p_{22}+p_{23}p_{32} & p_{22}p_{23}+p_{23}p_{33} \\
  	p_{32}p_{21}      & p_{32}p_{22}+p_{33}p_{32} & p_{32}p_{23}+p_{33}p_{33}
  \end{bmatrix}.
\end{equation}
It shows that after two steps of propagation, the probability of transition from $x_1$ to $x_3$ is no longer zero, instead it equals to the probability of transition from $x_1$ to $x_2$, $p_{12}$, in the first step multiplied by the probability of transition from $x_2$ to $x_3$, $p_{23}$, in the second step. Similar observation can also be found for other transitions among the data points.

It can be seen from the above that $P^t_{ij}$ sum all paths of length $t$ from point $x_i$ to point $x_j$. With increasing value of $t$, different scales of the data structure can be visualized. This is the diffusion process, during which the local connectivity is integrated to present the global connectivity of a data set. Besides, pathways built by long, low probability jumps are gradually replaced by short, high probability jumps. 

With the global geometric structure of a data set uncovered by the diffusion process described above, a diffusion metric can be defined for this structure. The metric measures the similarity of two points in the observed space as the connectivity between them, and is defined as
\begin{align}
	D_t(x_i,x_j)^2=&\sum_{u\in X}|p_t(x_i,u)-p_t(x_j,u)|^2\notag\\
	              =&\sum_k |P_{ik}^t-P_{kj}^t|^2.
\end{align}
In order to have a small diffusion distance between two points, there should be many high probability paths of length $t$ linking these two points. Since it sums over all possible paths, this algorithm is robust to noise perturbation. Besides, the path probabilities between $x_i$, $u$ and $u$, $x_j$ must be roughly equal. This happens when both $x_i$ and $x_j$ are well connected via $u$. However, calculating diffusion distances is computational expensive. Therefore, it is convenient to map data points into a Euclidean space, in which the diffusion distance becomes the Euclidean distance in this new diffusion space. 

A diffusion map maps coordinates between data and diffusion space by reorganizing data according to the diffusion metric. It preserves a data set's intrinsic geometry with data points mapped to a lower-dimensional structure. With the mapping
\begin{equation}
	y_i := \begin{bmatrix}
		p_t(x_i, x_{1}) \\
		p_t(x_i, x_{2}) \\
		\vdots \\
		p_t(x_i, x_{1N})
	\end{bmatrix}=P_{i\ast}^{\operatorname{T}},
\end{equation}
the Euclidean distance between two mapped points, $y_i$ and $y_j$, is
\begin{align}
	\lVert y_i-y_j\rVert_E^2=&\sum_{u\in X} |p_t(x_i,u)-p_t(x_j,u)|^2\notag\\
	                        =&\sum_k |P_{ik}^t-P_{kj}^t|^2\notag\\
	                        =&D_t(x_i,x_j)^2.
\end{align}

To reduce the dimensionality, we define the normalized diffusion matrix
\begin{equation}
	P=D^{-1}K,
\end{equation}
where $D$ is the diagonal matrix consisting of the row-sums of $K$. With the eigenvectors $\phi_i$ and $\lambda_i$ of $P$, the mapped points can be rewritten as
\begin{equation}
	y_i^\prime=\begin{bmatrix}
		\lambda_1^t\phi_1(i) \\
		\lambda_2^t\phi_2(i) \\
		\vdots \\
		\lambda_n^t\phi_n(i)
	\end{bmatrix},
\end{equation}
where $\phi_k(s)$ indicates the $s$th element of the $k$th eigenvector of $P$. The Euclidean distance between mapped points $y_i^\prime$ and $y_j^\prime$ is the diffusion distance. Dimension reduction is achieved by retaining the $m$ dimensions associated with the dominant eigenvectors.